---
title: "THESIS"
author: "Leah Mautner"
date: "2025-02-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(survey)
library(svydiags)
library(naniar)
library(stats)
library(jtools)
library(labelled)
library(knitr)
library(ggplot2)
library(lmtest)
library(broom)
library(gtsummary)
library(openxlsx)
```

# Import
```{r, echo= TRUE}
setwd("~/Desktop/GPH/Thesis")
load("NSDUH_2022.Rdata")
load("NSDUH_2023.Rdata")

NSDUH_2023 <- as.data.frame(puf2023_102124)
rm(puf2023_102124)
```

# Merge 
```{r, echo= TRUE}
colnames(NSDUH_2022) <- toupper(colnames(NSDUH_2022))
colnames(NSDUH_2023) <- toupper(colnames(NSDUH_2023))

NSDUH <- bind_rows(NSDUH_2022, NSDUH_2023)
head(NSDUH)
count(NSDUH)
```

#NAs
```{r, echo= TRUE}
NA_response <- c(85, 94, 97, 98, 99, ".")

NSDUH <- NSDUH %>%
  mutate(across(everything(), ~ ifelse(. %in% NA_response, NA, .)))
```

# Dependent Variable
```{r, echo= TRUE}
NSDUH <- NSDUH %>% mutate(
  Cost = as.integer(rowSums(across(c(MHNTCOST, MHNTINSCV, MHNTENFCV)), na.rm = TRUE) >= 1),
  Structural = as.integer(rowSums(across(c(MHNTWHER, MHNTNOFND, MHNTNOOPN, MHNTPROBS, MHNTTIME)), na.rm = TRUE) >= 1),
  Minimization = as.integer(rowSums(across(c(MHNTSTART, MHNTHNDL)), na.rm = TRUE) >= 1), 
  Perceived_Eff = as.integer(rowSums(across(c(MHNTNOHLP, MHNTCARE)), na.rm = TRUE) >= 1), 
  Stigma = as.integer(rowSums(across(c(MHNTPTHNK, MHNTCONSQ, MHNTFFLKE, MHNTFORCE, MHNTMEDS, MHNTPRIV)), na.rm = TRUE) >= 1)
)


NSDUH %>% summarise(Cost_n = sum(Cost, na.rm= TRUE), Structural_n= sum(Structural, na.rm= TRUE), Minimization_n= sum(Minimization, na.rm= TRUE), Perceived_n = sum(Perceived_Eff, na.rm= TRUE), Stigma_n = sum(Stigma, na.rm=TRUE))

NSDUH <- NSDUH %>%
  mutate(
    barrier_number = rowSums(across(c(
      MHNTCOST, MHNTINSCV, MHNTENFCV, MHNTWHER, MHNTNOFND, MHNTNOOPN, 
      MHNTPROBS, MHNTTIME, MHNTSTART, MHNTHNDL, MHNTNOHLP, MHNTCARE, 
      MHNTPTHNK, MHNTCONSQ, MHNTFFLKE, MHNTFORCE, MHNTMEDS, MHNTPRIV
    )), na.rm = TRUE)
  ) %>%
  mutate(barrier_number = replace_na(barrier_number, 0))  
table(NSDUH$barrier_number)

NSDUH <- NSDUH %>% mutate(barrier_groups = rowSums(across(c(Cost, Structural, Minimization, Perceived_Eff, Stigma)), na.rm=TRUE))
table(NSDUH$barrier_groups)
```


# Operationalize (When not imputed, N= 113 missing (~2%))
```{r, echo= TRUE}
#Insurance
NSDUH <- NSDUH %>%
  mutate(
    public_insur = ifelse(IRMEDICR == 1 | IRMCDCHP == 1 | IRCHMPUS == 1, 1, 0),
    private_insur = ifelse(IRPRVHLT == 1, 1, 0),
    other_insur = ifelse(IROTHHLT == 1, 1, 0),
    no_insur = ifelse(IRINSUR4 == 2, 1, 0)
  )

NSDUH <- NSDUH %>% mutate(Insurance = case_when(no_insur== 1 ~ 0, public_insur== 1 ~ 1, private_insur==1 ~2, other_insur==1~ 3))
table(NSDUH$Insurance)

NSDUH %>%
  mutate(
    Overlap = rowSums(across(c(public_insur, private_insur, other_insur, no_insur), ~ . == 1), na.rm = TRUE)
  ) %>%
  count(Overlap) 

table(NSDUH$public_insur, NSDUH$private_insur) # Overlap all comes from here, likely Medicare + private
table(NSDUH$public_insur, NSDUH$other_insur)    
table(NSDUH$private_insur, NSDUH$other_insur)  
table(NSDUH$no_insur, NSDUH$public_insur)       


NSDUH %>% count(is.na(Insurance))

#Distress
NSDUH$distress_level <- cut(NSDUH$KSSLR6MAX, 
                            breaks= c(0, 5, 13, 25),
                            labels= c("1", "2", "3"),
                            right= FALSE)
table(NSDUH$distress_level)

#Disability
NSDUH$disability_level <- cut(NSDUH$WHODASTOTSC, 
                            breaks= c(0, 8, 13, 25),
                            labels= c("1", "2", "3"),
                            right= FALSE)
table(NSDUH$disability_level)
```


# Rename
```{r, echo= TRUE}
colnames(NSDUH) <- toupper(colnames(NSDUH))

NSDUH <- NSDUH %>% rename(ID= QUESTID2, Cost= COST, Structural= STRUCTURAL, Minimization= MINIMIZATION, Low_perc_eff= PERCEIVED_EFF, Stigma= STIGMA, Age= CATAG6, Sex= IRSEX, Marital_status= IRMARIT, Education= EDUHIGHCAT, Race= NEWRACE2, Urbanicity= COUTYP4, Poverty= POVERTY3, Health_status= HEALTH, Employment_status= IRWRKSTAT18, Alcohol_Use= BNGDRKMON, Drug_Use= ILLMON, SUD= UD5ILALANY, Whodas_raw= WHODASTOTSC, Kessler_raw= KSSLR6MAX, Insurance=INSURANCE, Distress=DISTRESS_LEVEL, Disability=DISABILITY_LEVEL, Barrier_total= BARRIER_NUMBER, Barrier_grouptotal = BARRIER_GROUPS, Subset_Utilization= MHTRTPY, Subset_UnmetNeed= MHTSKTHPY, Subset_Age= CATAGE, Income= INCOME)
```

# Variable Types and Labels
```{r, echo= TRUE}
NSDUH <- NSDUH %>% mutate(across(c(Cost, Structural, Minimization, Low_perc_eff, Stigma), as.integer))

NSDUH <- NSDUH %>% mutate(across(c(Whodas_raw, Kessler_raw), as.numeric))  

NSDUH <- NSDUH %>% mutate(
  Age= factor(Age,
              levels= c(1, 2, 3, 4, 5, 6),
              labels= c("12-17", "18-25", "26-34", "35-49", "50+", "50+")),
  Sex= factor(Sex, 
              levels= c(1,2),
              labels= c("Male", "Female")), 
  Marital_status= factor(Marital_status, 
                         levels= c(1,2,3,4),
                         labels= c("Married", "Widowed/Divorced/Separated", "Widowed/Divorced/Separated", "Never Married")),
  Education= factor(Education,
                    levels= c(1,2,3,4,5),
                    labels= c("Less than HS", "HS Grad", "Some College/Assoc Degree", "Colleg Grad", "12-17")),
  Race= factor(Race,
               levels= c(1,2,3,4,5,6,7),
               labels= c("Non-Hisp White", "Non-Hisp Black/Afr Am", "Non-Hisp Native Am/AK Native/Native HI/Other PI", "Non-Hisp Native Am/AK Native/Native HI/Other PI", "Non-Hisp Asian", "Non-Hisp Multi Race", "Hispanic")),
  Urbanicity= factor(Urbanicity,
                     levels= c(1,2,3),
                     labels= c("Large Metro", "Small Metro", "Nonmetro")),
  Poverty= factor(Poverty,
                  levels= c(1,2,3),
                  labels= c("Living in Poverty", "Income Up to 2x FPL", "Income More than 2x FPL")),
  Income= factor(Income,
                 levels= c(1, 2, 3, 4),
                 labels= c("Less than $20,000", "$20,000-$49,000", "50,000-$75,000", "$75,000 or More")),
  Health_status= factor(Health_status,
                        levels= c(1,2,3,4,5),
                        labels= c("Excellent", "Very Good", "Good", "Fair", "Poor")),
  Employment_status= factor(Employment_status,
                            levels= c(1,2,3,4),
                            labels= c("Full time", "Part time", "Unemployed", "Other")),
  SUD= factor(SUD,
              levels= c(0,1),
              labels= c("No SUD", "SUD")),
  Drug_Use= factor(Drug_Use, 
                   levels= c(0,1),
                   labels= c("No Past Month Use", "Past Month Use")),
  Alcohol_Use= factor(Alcohol_Use, 
                      levels= c(0,1),
                      labels= c("No Past Month Binge Drinking", "Past Month Binge Drinking")),
  Insurance= factor(Insurance, 
                    levels= c(0,1,2,3),
                    labels= c("No Insurance", "Public Insurance", "Private Insurance", "Other Insurance")),
  Distress= factor(Distress,
                   levels= c(1,2,3),
                   labels= c("Low PD", "Moderate PD", "Severe PD")),
  Disability= factor(Disability,
                     levels= c(1,2,3),
                     labels= c("Low Disability", "Moderate Disability", "High Disability")),
)

NSDUH$Age <- droplevels(NSDUH$Age)
NSDUH$Marital_status <- droplevels(NSDUH$Marital_status)
NSDUH$Race <- droplevels(NSDUH$Race)

NSDUH <- NSDUH %>%
  mutate(
    Race = as.character(Race),                  
    Race = ifelse(Race == "Non-Hisp Native Am/AK Native/Native HI/Other PI", 
                  "Non-Hisp Multi Race", Race),  
    Race = factor(Race)                        
  ) %>%
  mutate(
    Race = droplevels(Race)                     
  )
table(NSDUH$Race)
```


# Check for collinearity 
```{r, echo= TRUE}
# Cross-Tab 
dv_data <- NSDUH_sample %>%
  dplyr::select(Cost, Structural, Minimization, Low_perc_eff, Stigma) %>%
  mutate(across(everything(), as.numeric)) 

crosstab_counts <- t(as.matrix(dv_data)) %*% as.matrix(dv_data)

crosstab_percent <- round(100 * crosstab_counts / nrow(NSDUH_sample), 2)

crosstab_combined <- as.data.frame(crosstab_counts)
crosstab_combined[] <- paste0(crosstab_counts, " (", crosstab_percent, "%)")

kable(crosstab_combined, format = "markdown", caption = "Barrier Group Overlap (Counts & Percentages)")


Cor_1 <- NSDUH %>% dplyr::filter(Subset_UnmetNeed == 1 & Subset_Age != 1) %>% dplyr::select(Cost, Structural, Minimization, Low_perc_eff, Stigma) %>% cor(use="complete.obs")

print(Cor_1)
Cor_1 <- as.data.frame(Cor_1)
Cor_1_Round <- round(Cor_1, 4)
write.csv(Cor_1_Round, "correlation_matrix.csv")

#significance
vars <- colnames(Cor_1)
n <- length(vars)
p_value <- matrix(NA, n, n)

for(i in 1:(n-1)) {
  for(j in (i+1):n) {
    var1 <- vars[i]
    var2 <- vars[j]
test_result <- cor.test(Cor_1[,i], Cor_1[,j], method = "pearson")
    p_value[i, j] <- test_result$p.value
    p_value[j, i] <- test_result$p.value  
  }
}

p_valuedf <- as.data.frame(p_value)
colnames(p_valuedf) <- rownames(p_valuedf) <- vars

print(p_valuedf) #no significant correlations 
```

```{r}
library(polycor)
library(knitr)

# Compute tetrachoric correlation matrix
tetrachoric_corr_matrix <- hetcor(NSDUH_sample[, dv_groups], std.err = TRUE)

# Extract correlation coefficients
cor_matrix <- round(tetrachoric_corr_matrix$correlations, 3)

# Extract standard errors
se_matrix <- round(tetrachoric_corr_matrix$std.errors, 3)

# Compute 95% Confidence Intervals
ci_lower <- cor_matrix - 1.96 * se_matrix
ci_upper <- cor_matrix + 1.96 * se_matrix

# Mark significance based on confidence intervals
p_values <- ifelse(ci_lower * ci_upper > 0, "< 0.05", "≥ 0.05")

# Format matrix with coefficients and p-values
formatted_matrix <- matrix("", nrow = length(dv_groups), ncol = length(dv_groups))
rownames(formatted_matrix) <- colnames(formatted_matrix) <- dv_groups

for (i in 1:length(dv_groups)) {
  for (j in 1:length(dv_groups)) {
    if (i == j) {
      formatted_matrix[i, j] <- "1 (NA)"
    } else {
      formatted_matrix[i, j] <- paste0(cor_matrix[i, j], " (", p_values[i, j], ")")
    }
  }
}

# Print the table
kable(formatted_matrix, caption = "Unweighted Tetrachoric Correlation Matrix with Confidence Intervals for Significance")


```


# Create DF
```{r, echo= TRUE}
NSDUH <- NSDUH %>% dplyr::select(ID, VEREP, VESTR_C, ANALWT2_C, Subset_Utilization, Subset_UnmetNeed, Subset_Age, Cost, Structural, Minimization, Low_perc_eff, Stigma, Whodas_raw, Kessler_raw, Barrier_total, Barrier_grouptotal, Age, Sex, Marital_status, Education, Race, Urbanicity, Poverty, Health_status, Employment_status, SUD, Insurance, Distress, Disability, Income, Drug_Use, Alcohol_Use)
count(NSDUH)

NSDUH %>%
  dplyr::filter(Subset_UnmetNeed == 1 & Subset_Age != 1) %>%
  summarise(across(everything(), ~ sum(is.na(.)))) #Insurance NA= 113, Poverty NA= 3, Health Status NA= 2

NSDUH %>% dplyr::filter(Subset_UnmetNeed == 1 & Subset_Age != 1) %>% filter(is.na(Insurance)) %>% pull(Distress) %>% table() #N=113 missing for age, 74 of which are severe PD when not using imputed*

NSDUH <- NSDUH %>% drop_na(Poverty) %>% drop_na(Health_status)

NSDUH %>% dplyr::filter(Subset_Age != 1) %>% pull(Subset_Utilization) %>% table()

NSDUH_sample <- NSDUH %>%
  dplyr::filter(Subset_UnmetNeed == 1 & Subset_Age != 1)


str(NSDUH)
```

# Closer look 
```{r, echo= TRUE}
table(NSDUH_sample$Stigma, NSDUH_sample$Minimization)
table(NSDUH_sample$Stigma, NSDUH_sample$Low_perc_eff)
table(NSDUH_sample$Minimization, NSDUH_sample$Low_perc_eff)
table(NSDUH_sample$Cost, NSDUH_sample$Structural)

table(NSDUH_sample$Stigma, NSDUH_sample$Race)
```

# Analytic Sample 
```{r, echo= TRUE}
NSDUH %>% filter(Subset_Age != 1) %>% pull(Subset_Utilization) %>% table()

NSDUH %>% filter(Subset_Age != 1 & Subset_Utilization == 0) %>% pull(Subset_UnmetNeed) %>% table()
```

# Missing Observations 
```{r, echo= TRUE}
Missing <- NSDUH %>% filter(Subset_Age != 1)
Missing <- Missing %>% filter(Subset_Utilization == 0 & is.na(Subset_UnmetNeed))

total_n <- nrow(Missing)

bind_rows(
  as.data.frame(table(Missing$Age)) %>% mutate(Variable = "Age"),
  as.data.frame(table(Missing$Sex)) %>% mutate(Variable = "Sex"),
  as.data.frame(table(Missing$Marital_status)) %>% mutate(Variable = "Marital_status"),
  as.data.frame(table(Missing$Education)) %>% mutate(Variable = "Education"),
  as.data.frame(table(Missing$Race)) %>% mutate(Variable = "Race"),
  as.data.frame(table(Missing$Urbanicity)) %>% mutate(Variable = "Urbanicity"),
  as.data.frame(table(Missing$Poverty)) %>% mutate(Variable = "Poverty"),
  as.data.frame(table(Missing$Health_status)) %>% mutate(Variable = "Health_status"),
  as.data.frame(table(Missing$Employment_status)) %>% mutate(Variable = "Employment_status"),
  as.data.frame(table(Missing$SUD)) %>% mutate(Variable = "SUD"),
  as.data.frame(table(Missing$Insurance)) %>% mutate(Variable = "Insurance"),
  as.data.frame(table(Missing$Disability)) %>% mutate(Variable = "Disability"),
    as.data.frame(table(Missing$Distress)) %>% mutate(Variable = "Distress")
) %>%
  rename(Level = Var1, Count = Freq) %>%
  mutate(Proportion = round((Count / total_n) * 100, 2)) #Seems relatively normally distributed
```

# Export to STATA
```{r}
library(haven)
library(dplyr)
library(labelled)

categorical_vars <- c("Age", "Sex", "Marital_status", "Education", "Race", "Urbanicity",
                      "Poverty", "Health_status", "Employment_status", "SUD", "Insurance",
                      "Disability", "Education", "Income", "Drug_Use", "Alcohol_Use", "Distress")

NSDUH <- NSDUH %>%
  mutate(across(all_of(categorical_vars), as.factor)) 

for (var in categorical_vars) {
  NSDUH[[var]] <- labelled::to_factor(NSDUH[[var]])
}

write_dta(NSDUH, "NSDUH.dta", version = 14)
```


# Survey Weights
```{r, echo= TRUE}
NSDUH <- NSDUH %>% mutate(across(c(Cost, Structural, Minimization, Low_perc_eff, Stigma), as.factor))

NSDUH <- NSDUH %>% mutate(across(c(VEREP, VESTR_C), as.factor))
class(NSDUH$ANALWT2_C)

NSDUH$ANALWT2_Adjusted <- NSDUH$ANALWT2_C/2

NSDUH_weights <- svydesign(
  id= ~ VEREP,
  strata= ~ VESTR_C,
  weights= ~ ANALWT2_Adjusted,
  data= NSDUH,
  nest= TRUE
)
summary(NSDUH_weights)

NSDUH_subset <- subset(NSDUH_weights, Subset_UnmetNeed == 1 & Subset_Age != 1)
head(NSDUH_subset)
str(NSDUH_subset)
nrow(NSDUH_subset$variables)

NSDUH_subset <- update(NSDUH_subset, Age = droplevels(as.factor(Age)))
NSDUH_subset <- update(NSDUH_subset, Education = droplevels(as.factor(Education)))
```

# Weighted DV Collinearity 
```{r}
library(survey)

barrier_vars <- c("Cost", "Structural", "Minimization", "Low_perc_eff", "Stigma") 


for (var in barrier_vars) {
  NSDUH_subset$variables[[var]] <- as.numeric(NSDUH_subset$variables[[var]])
}


cov_matrix <- svyvar(as.formula(paste("~", paste(barrier_vars, collapse = " + "))), design = NSDUH_subset)
cor_matrix <- cov2cor(as.matrix(cov_matrix))


se_matrix <- sqrt(diag(cov_matrix))
p_values <- 2 * (1 - pnorm(abs(cor_matrix / outer(se_matrix, se_matrix, "*"))))

cor_matrix <- round(cor_matrix, 3)
p_values <- round(p_values, 3)

cor_p_matrix <- matrix("", nrow=length(barrier_vars), ncol=length(barrier_vars))
rownames(cor_p_matrix) <- colnames(cor_p_matrix) <- barrier_vars

for (i in 1:length(barrier_vars)) {
  for (j in 1:length(barrier_vars)) {
    cor_p_matrix[i, j] <- paste0(cor_matrix[i, j], " (", p_values[i, j], ")")  
  }
}

kable(cor_p_matrix, caption = "Survey-Weighted Correlation Matrix with P-values")
table(NSDUH_subset$Cost, NSDUH_subset$Structural)
```

# Tetrachoric Correlation Matrix (binary)
```{r}
library(polycor)

compute_tetrachoric_correlations <- function(design, dv_groups) {
  design$variables[dv_groups] <- lapply(design$variables[dv_groups], function(x) as.numeric(x == TRUE))

  tetrachoric_results <- hetcor(design$variables[, dv_groups], std.err = TRUE)

  cor_matrix <- round(tetrachoric_results$correlations, 3)

  return(list(cor_matrix = cor_matrix))
}

tetra_results <- compute_tetrachoric_correlations(NSDUH_subset, dv_groups)
view(tetra_results)
```


# Weighted Univariable
## Continuous 
```{r, echo= TRUE}
svymean(~Kessler_raw, design= NSDUH_subset)
svyquantile(~Kessler_raw, design= NSDUH_subset, quantiles=c(0.25, 0.5, 0.75))
sqrt(svyvar(~Kessler_raw, design = NSDUH_subset))

svymean(~Barrier_total, design= NSDUH_subset)
svyquantile(~Barrier_total, design= NSDUH_subset, quantiles=c(0.25, 0.5, 0.75))
sqrt(svyvar(~Barrier_total, design = NSDUH_subset))

NSDUH_subset$Barrier_grouptotal <- as.integer(NSDUH_subset$Barrier_grouptotal)
table(NSDUH_subset$Barrier_grouptotal)

svymean(~ Barrier_grouptotal, design= NSDUH_subset)
svyquantile(~ Barrier_grouptotal, design= NSDUH_subset, quantiles=c(0.25, 0.5, 0.75))
sqrt(svyvar(~Barrier_grouptotal, design = NSDUH_subset))
```

## Categorical
```{r, echo= TRUE}
categorical_vars <- c("Cost", "Structural", "Minimization", "Low_perc_eff", "Stigma", "Age", "Sex", "Marital_status", "Education", "Race", "Urbanicity", "Poverty", "Health_status", "Employment_status", "SUD", "Insurance", "Distress", "Disability")

# Prop
calculate_proportions <- function(categorical_vars, NSDUH_subset) {
  results <- list()
  
  for (var in categorical_vars) {
    formula_var <- as.formula(paste0("~", var))
    
    svy_props <- svymean(formula_var, NSDUH_subset)
    
    results[[var]] <- data.frame(
      Category = names(svy_props),
      Proportion = round(as.numeric(svy_props) * 100, 2)  
    )
  }
  
  return(results)
}

proportions <- calculate_proportions(categorical_vars, NSDUH_subset)
print(proportions)

# 95% CI
calculate_confidence_intervals <- function(categorical_vars, NSDUH_subset) {
  results <- list()
  
  for (var in categorical_vars) {
    formula_var <- as.formula(paste0("~", var))
    
    # Compute weighted proportions with confidence intervals
    svy_props <- svymean(formula_var, NSDUH_subset, vartype = "ci")
    
    # Store formatted results
    results[[var]] <- data.frame(
      Category = names(svy_props),
      Lower_CI = round(confint(svy_props)[, 1] * 100, 2),  # Lower bound
      Upper_CI = round(confint(svy_props)[, 2] * 100, 2)   # Upper bound
    )
  }
  
  return(results)
}

conf_intervals <- calculate_confidence_intervals(categorical_vars, NSDUH_subset)
print(conf_intervals)

prop <- svymean(~factor(Race), design = NSDUH_subset, vartype = "ci")
confint(prop)
```


# Unweighted Univariable
## Continuous 
```{r, echo= TRUE}
mean(NSDUH_sample$Kessler_raw)
quantile(NSDUH_sample$Kessler_raw, probs=c(0.25, 0.5, 0.75))
sd(NSDUH_sample$Kessler_raw)
ggplot(data = NSDUH_sample) + geom_boxplot(aes(x="", y= Kessler_raw))

mean(NSDUH_sample$Barrier_total)
quantile(NSDUH_sample$Barrier_total, probs=c(0.25, 0.5, 0.75))
sd(NSDUH_sample$Barrier_total)
ggplot(data = NSDUH_sample) + geom_boxplot(aes(x="", y= Barrier_total))

mean(NSDUH_sample$Barrier_grouptotal)
quantile(NSDUH_sample$Barrier_grouptotal, probs=c(0.25, 0.5, 0.75))
sd(NSDUH_sample$Barrier_grouptotal)
ggplot(data = NSDUH_sample) + geom_boxplot(aes(x="", y= Barrier_grouptotal))
```


## Categorical
```{r, echo= TRUE}
unweighted_categorical_univariable <- function(categorical_vars, NSDUH_sample) {
  results_unweighted <- list()
  
  for(var in categorical_vars) {
    raw_count <- table(NSDUH_sample[[var]])
    raw_prop <- prop.table(raw_count) * 100
    
    summary_tab <- data.frame(Category_=names(raw_count), Count=as.numeric(raw_count), Percentage_=round(raw_prop, 3))
  
results_unweighted[[var]] <- summary_tab
  }
  return(results_unweighted)
}

univariable_unweighted <- unweighted_categorical_univariable(categorical_vars, NSDUH_sample)

lapply(univariable_unweighted, function(NSDUH_sample) kable(NSDUH_sample))

NSDUH_sample %>% count(Income) %>% mutate(pct = 100 * n / sum(n)) %>% kable()
NSDUH_sample %>% count(Drug_Use) %>% mutate(pct = 100 * n / sum(n)) %>% kable()
```

# Normality Test 
```{r, echo= TRUE}
library(nortest)
ad.test(NSDUH_sample$Kessler_raw)
ad.test(NSDUH_sample$Barrier_total)
```

# Weighted Bivariable for IV 
## Categorical
```{r}
distress_cat_bivariable <- svyby(
  ~ Cost + Structural + Minimization + Low_perc_eff + Stigma + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability,
  ~ Distress,
  NSDUH_subset,
  svymean,
  vartype= "ci"
) 

distress_cat_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()
```
## Continuous 
```{r}
distress_cont_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Distress,
  NSDUH_subset,
  svymean,
  sd=sqrt(svyvar()),
  vartype= NULL
) 

distress_cont_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

distress_contin_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Distress,
  NSDUH_subset,
  svyquantile,
  quantiles=c(0.25,0.50,0.75),
  vartype= NULL
)

distress_contin_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()
```

## Chi-sq (CHECK W SOMEONE)
```{r}
NSDUH_subset$Age <- droplevels(NSDUH_subset$Age)
NSDUH_subset$Education <- droplevels(NSDUH_subset$Education)

categorical_var <- c("Cost", "Structural", "Minimization", "Low_perc_eff", "Stigma", "Age", "Sex", "Marital_status", "Education", "Race", "Urbanicity", "Poverty", "Health_status", "Employment_status", "SUD", "Insurance", "Disability")

chi_square <- function(var) {
  test <- svychisq(as.formula(paste("~", var, "+ Distress")), design = NSDUH_subset)
  result <- data.frame(
    Variable = var,
    Chi_Square_Stat = round(test$statistic, 3),
    P_Value = ifelse(test$p.value < 0.001, "<0.001", round(test$p.value, 3))
  )
  return(result)
}

results <- do.call(rbind, lapply(categorical_var, chi_square))
kable(results)
```

## Continuous 
```{r}
wald_svy <- function(var) {
  model <- svyglm(as.formula(paste(var, "~ Distress")), design = NSDUH_subset)
  
  test <- regTermTest(model, ~ Distress)
  
  result <- data.frame(
    Variable = var,
    Wald_Statistic = round(test$Ftest, 3),
    P_Value = ifelse(test$p < 0.001, "<0.001", round(test$p, 3))
  )
  
  return(result)
}

kable(wald_svy("Kessler_raw"))
kable(wald_svy("Barrier_total"))
kable(wald_svy("Barrier_grouptotal"))
```

# Weighted Bivariable for DV 
```{r}
# Cost 
# Categorical
cost_cat_bivariable <- svyby(
  ~ Structural + Minimization + Low_perc_eff + Stigma + Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability,
  ~ Cost,
  NSDUH_subset,
  svymean,
  vartype= "ci"
) 

cost_cat_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Continuous
cost_cont_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Cost,
  NSDUH_subset,
  svymean,
  sd=sqrt(svyvar()),
  vartype= NULL
) 

cost_cont_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

cost_contin_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Cost,
  NSDUH_subset,
  svyquantile,
  quantiles=c(0.25,0.50,0.75),
  vartype= NULL
)

cost_contin_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Chi-sq
categorical_var1 <- c("Structural", "Minimization", "Low_perc_eff", "Stigma", "Age", "Sex", "Marital_status", "Education", "Race", "Urbanicity", "Poverty", "Health_status", "Employment_status", "SUD", "Insurance", "Distress", "Disability")

chi_square1 <- function(var) {
  test <- svychisq(as.formula(paste("~", var, "+ Cost")), design = NSDUH_subset)
  result1 <- data.frame(
    Variable = var,
    Chi_Square_Stat = round(test$statistic, 3),
    P_Value = ifelse(test$p.value < 0.001, "<0.001", round(test$p.value, 3))
  )
  return(result1)
}

results1 <- do.call(rbind, lapply(categorical_var1, chi_square1))
kable(results1)

# Wald
wald_svy1 <- function(var) {
  model <- svyglm(as.formula(paste(var, "~ Cost")), design = NSDUH_subset)
  
  test <- regTermTest(model, ~ Cost)
  
  result1 <- data.frame(
    Variable = var,
    Wald_Statistic = round(test$Ftest, 3),
    P_Value = ifelse(test$p < 0.001, "<0.001", round(test$p, 3))
  )
  
  return(result1)
}

kable(wald_svy1("Kessler_raw"))
kable(wald_svy1("Barrier_total"))
kable(wald_svy1("Barrier_grouptotal"))
```

```{r}
# Structural 
# Categorical
structural_cat_bivariable <- svyby(
  ~ Cost + Minimization + Low_perc_eff + Stigma + Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability,
  ~ Structural,
  NSDUH_subset,
  svymean,
  vartype= "ci"
) 

structural_cat_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Continuous
structural_cont_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Structural,
  NSDUH_subset,
  svymean,
  sd=sqrt(svyvar()),
  vartype= NULL
) 

structural_cont_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

structural_contin_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Structural,
  NSDUH_subset,
  svyquantile,
  quantiles=c(0.25,0.50,0.75),
  vartype= NULL
)

structural_contin_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Chi-sq
categorical_var2 <- c("Cost", "Minimization", "Low_perc_eff", "Stigma", "Age", "Sex", "Marital_status", "Education", "Race", "Urbanicity", "Poverty", "Health_status", "Employment_status", "SUD", "Insurance", "Distress", "Disability")

chi_square2 <- function(var) {
  test <- svychisq(as.formula(paste("~", var, "+ Structural")), design = NSDUH_subset)
  result2 <- data.frame(
    Variable = var,
    Chi_Square_Stat = round(test$statistic, 3),
    P_Value = ifelse(test$p.value < 0.001, "<0.001", round(test$p.value, 3))
  )
  return(result2)
}

results2 <- do.call(rbind, lapply(categorical_var2, chi_square2))
kable(results2)

# Wald
wald_svy2 <- function(var) {
  model <- svyglm(as.formula(paste(var, "~ Structural")), design = NSDUH_subset)
  
  test <- regTermTest(model, ~ Structural)
  
  result2 <- data.frame(
    Variable = var,
    Wald_Statistic = round(test$Ftest, 3),
    P_Value = ifelse(test$p < 0.001, "<0.001", round(test$p, 3))
  )
  
  return(result2)
}

kable(wald_svy2("Kessler_raw"))
kable(wald_svy2("Barrier_total"))
kable(wald_svy2("Barrier_grouptotal"))

#NAs
svychisq(~ Age + Stigma, design = NSDUH_subset)
svychisq(~ Education + Stigma, design = NSDUH_subset)
```

```{r}
# Minimization
# Categorical
min_cat_bivariable <- svyby(
  ~ Cost + Structural + Low_perc_eff + Stigma + Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability,
  ~ Minimization,
  NSDUH_subset,
  svymean,
  vartype= "ci"
) 

min_cat_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Continuous
min_cont_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Minimization,
  NSDUH_subset,
  svymean,
  sd=sqrt(svyvar()),
  vartype= NULL
) 

min_cont_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

min_contin_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Minimization,
  NSDUH_subset,
  svyquantile,
  quantiles=c(0.25,0.50,0.75),
  vartype= NULL
)

min_contin_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Chi-sq
categorical_var3 <- c("Cost", "Structural", "Low_perc_eff", "Stigma", "Age", "Sex", "Marital_status", "Education", "Race", "Urbanicity", "Poverty", "Health_status", "Employment_status", "SUD", "Insurance", "Distress", "Disability")

chi_square3 <- function(var) {
  test <- svychisq(as.formula(paste("~", var, "+ Minimization")), design = NSDUH_subset)
  result3 <- data.frame(
    Variable = var,
    Chi_Square_Stat = round(test$statistic, 3),
    P_Value = ifelse(test$p.value < 0.001, "<0.001", round(test$p.value, 3))
  )
  return(result3)
}

results3 <- do.call(rbind, lapply(categorical_var3, chi_square3))
kable(results3)

# Wald
wald_svy3 <- function(var) {
  model <- svyglm(as.formula(paste(var, "~ Minimization")), design = NSDUH_subset)
  
  test <- regTermTest(model, ~ Minimization)
  
  result3 <- data.frame(
    Variable = var,
    Wald_Statistic = round(test$Ftest, 3),
    P_Value = ifelse(test$p < 0.001, "<0.001", round(test$p, 3))
  )
  
  return(result3)
}

kable(wald_svy3("Kessler_raw"))
kable(wald_svy3("Barrier_total"))
kable(wald_svy3("Barrier_grouptotal"))
```

```{r}
# Low Perceived Effectiveness 
# Categorical
loweff_cat_bivariable <- svyby(
  ~ Structural + Minimization + Cost + Stigma + Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability,
  ~ Low_perc_eff,
  NSDUH_subset,
  svymean,
  vartype= "ci"
) 

loweff_cat_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Continuous
loweff_cont_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Low_perc_eff,
  NSDUH_subset,
  svymean,
  sd=sqrt(svyvar()),
  vartype= NULL
) 

loweff_cont_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

loweff_contin_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Low_perc_eff,
  NSDUH_subset,
  svyquantile,
  quantiles=c(0.25,0.50,0.75),
  vartype= NULL
)

loweff_contin_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Chi-sq
categorical_var4 <- c("Cost", "Structural", "Minimization", "Stigma", "Age", "Sex", "Marital_status", "Education", "Race", "Urbanicity", "Poverty", "Health_status", "Employment_status", "SUD", "Insurance", "Distress", "Disability")

chi_square4 <- function(var) {
  test <- svychisq(as.formula(paste("~", var, "+ Low_perc_eff")), design = NSDUH_subset)
  result4 <- data.frame(
    Variable = var,
    Chi_Square_Stat = round(test$statistic, 3),
    P_Value = ifelse(test$p.value < 0.001, "<0.001", round(test$p.value, 3))
  )
  return(result4)
}

results4 <- do.call(rbind, lapply(categorical_var4, chi_square4))
kable(results4)

# Wald
wald_svy4 <- function(var) {
  model <- svyglm(as.formula(paste(var, "~ Low_perc_eff")), design = NSDUH_subset)
  
  test <- regTermTest(model, ~ Low_perc_eff)
  
  result4 <- data.frame(
    Variable = var,
    Wald_Statistic = round(test$Ftest, 3),
    P_Value = ifelse(test$p < 0.001, "<0.001", round(test$p, 3))
  )
  
  return(result4)
}

kable(wald_svy4("Kessler_raw"))
kable(wald_svy4("Barrier_total"))
kable(wald_svy4("Barrier_grouptotal"))
```

```{r}
# Stigma 
# Categorical
stigma_cat_bivariable <- svyby(
  ~ Structural + Minimization + Low_perc_eff + Cost + Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability,
  ~ Stigma,
  NSDUH_subset,
  svymean,
  vartype= "ci"
) 

stigma_cat_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Continuous 
stigma_cont_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Stigma,
  NSDUH_subset,
  svymean,
  sd=sqrt(svyvar()),
  vartype= NULL
) 

stigma_cont_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

stigma_contin_bivariable <- svyby(
  ~ Kessler_raw + Barrier_total + Barrier_grouptotal,
  ~ Stigma,
  NSDUH_subset,
  svyquantile,
  quantiles=c(0.25,0.50,0.75),
  vartype= NULL
)

stigma_contin_bivariable %>% mutate(across(where(is.numeric), round, 3)) %>% kable()

# Chi-sq
categorical_var5 <- c("Cost", "Structural", "Low_perc_eff", "Minimization", "Age", "Sex", "Marital_status", "Education", "Race", "Urbanicity", "Poverty", "Health_status", "Employment_status", "SUD", "Insurance", "Distress", "Disability")

chi_square5 <- function(var) {
  test <- svychisq(as.formula(paste("~", var, "+ Stigma")), design = NSDUH_subset)
  result5 <- data.frame(
    Variable = var,
    Chi_Square_Stat = round(test$statistic, 3),
    P_Value = ifelse(test$p.value < 0.001, "<0.001", round(test$p.value, 3))
  )
  return(result5)
}

results5 <- do.call(rbind, lapply(categorical_var5, chi_square5))
kable(results5)

# Wald
wald_svy5 <- function(var) {
  model <- svyglm(as.formula(paste(var, "~ Stigma")), design = NSDUH_subset)
  
  test <- regTermTest(model, ~ Stigma)
  
  result5 <- data.frame(
    Variable = var,
    Wald_Statistic = round(test$Ftest, 3),
    P_Value = ifelse(test$p < 0.001, "<0.001", round(test$p, 3))
  )
  
  return(result5)
}

kable(wald_svy5("Kessler_raw"))
kable(wald_svy5("Barrier_total"))
kable(wald_svy5("Barrier_grouptotal"))
```

# Row Totals Bivariable Table 2.0
```{r}
library(survey)
library(dplyr)
library(tidyr)
library(knitr)

categorical_var <- c("Age", "Sex", "Marital_status", "Education", "Race", "Urbanicity", "Poverty", "Income", "Health_status", "Employment_status", "SUD", "Drug_Use", "Alcohol_Use", "Insurance", "Distress", "Disability")

dv_groups <- c("Cost", "Structural", "Minimization", "Low_perc_eff", "Stigma")

svyby(~Cost, ~Race, NSDUH_subset, svyciprop, method = "logit", vartype = "ci")
svyby(~Structural, ~Race, NSDUH_subset, svyciprop, method = "logit", vartype = "ci")
svyby(~Minimization, ~Race, NSDUH_subset, svyciprop, method = "logit", vartype = "ci")
svyby(~Low_perc_eff, ~Race, NSDUH_subset, svyciprop, method = "logit", vartype = "ci")
svyby(~Stigma, ~Race, NSDUH_subset, svyciprop, method = "logit", vartype = "ci")

barriers_by_distress <- function(barrier) {
  prop_table <- svyby(as.formula(paste0("~", barrier)), ~Distress, NSDUH_subset, svyciprop, method = "logit", vartype = "ci")
  chi_test <- svychisq(as.formula(paste0("~", barrier, " + Distress")), NSDUH_subset)
   p_value <- round(chi_test$p.value, 3)
  list(Proportions = prop_table, Chi_Square = chi_test)
}

results_ <- lapply(dv_groups, barriers_by_distress)

results_
```

```{r}
distress_race <- svyby(
  ~ Distress, 
  ~ Race, 
  NSDUH_subset, 
  svymean, 
  na.rm = TRUE, 
  vartype = c("ci") # Directly computes confidence intervals
)

chi_test <- svychisq(~ Distress + Race, NSDUH_subset)

kable(distress_race, digits = 3, caption = "Proportion of Distress Levels by Race/Ethnicity with 95% CI")
print(chi_test)

#Adding new var
svymean(~ Income, NSDUH_subset)
confint(svymean(~ Income, NSDUH_subset))
svymean(~ Drug_Use, NSDUH_subset)
confint(svymean(~ Drug_Use, NSDUH_subset))
svymean(~ Alcohol_Use, NSDUH_subset)
confint(svymean(~ Alcohol_Use, NSDUH_subset))

```

# Chi-sq rows
```{r}
barriers_by_cov <- function(barrier) {
  svyby(
    ~ Drug_Use + Income + Alcohol_Use,  
    ~ dv_groups,  
    NSDUH_subset,
    svymean,
    vartype = "ci"
  )
}

barrier_results <- lapply(dv_groups, barriers_by_cov)

barrier_results

```


# Correlation among predictors 
```{r}
NSDUH_long <- NSDUH_long %>%
  mutate(across(c(Distress, Age, Sex, Marital_status, Race, Urbanicity, 
                  Poverty, Health_status, Employment_status, SUD, Insurance, 
                  Disability, Income, Drug_Use, Alcohol_Use, Barrier_Present), as.numeric))

numeric_df <- NSDUH_long %>%
  dplyr::select(Distress, Age, Sex, Marital_status, Race, Urbanicity, 
         Poverty, Health_status, Employment_status, SUD, Insurance, 
         Disability, Income, Drug_Use, Alcohol_Use) %>%
  mutate_all(as.numeric)

cor_matrix <- cor(numeric_df, use = "pairwise.complete.obs")
caret::findCorrelation(cor_matrix, cutoff = 0.9, names = TRUE)
```


```{r}
NSDUH_weighted <- as.data.frame(NSDUH_subset$variables)
```

# Logistic Regressions 

```{r}
barriers <- c("Cost", "Structural", "Minimization", "Low_perc_eff", "Stigma")

for (var in barriers) {
  NSDUH_subset[[var]] <- ifelse(as.character(NSDUH_subset[[var]]) == "Yes", 1,
                                ifelse(as.character(NSDUH_subset[[var]]) == "No", 0, NA))
}
```


```{r}
#Outcome ~ Predictor
model1_cost <- svyglm(Cost ~ Distress, design= NSDUH_subset, family=quasibinomial())
model1_struc <- svyglm(Structural ~ Distress, design= NSDUH_subset, family=quasibinomial())
model1_min <- svyglm(Minimization ~ Distress, design= NSDUH_subset, family=quasibinomial())
model1_lowp <- svyglm(Low_perc_eff ~ Distress, design= NSDUH_subset, family=quasibinomial())
model1_stigma <- svyglm(Stigma ~ Distress, design= NSDUH_subset, family=quasibinomial())

wb <- createWorkbook()

model_list <- list(
  Cost         = model1_cost,
  Structural   = model1_struc,
  Minimization = model1_min,
  Low_Perc_Eff = model1_lowp,
  Stigma       = model1_stigma
)

for (name in names(model_list)) {
  sheet_data <- tidy(model_list[[name]],
                     exponentiate = TRUE, conf.int = TRUE) %>%
    select(term, estimate, conf.low, conf.high, p.value) %>%
    rename(
      OR = estimate,
      `95% CI Lower` = conf.low,
      `95% CI Upper` = conf.high,
      `P-value` = p.value
    )
  
  addWorksheet(wb, name)
  writeData(wb, sheet = name, sheet_data)
}

saveWorkbook(wb, "model1_output.xlsx", overwrite = TRUE)
```

```{r}
NSDUH_subset$variables$Race <- relevel(NSDUH_subset$variables$Race, ref = "Non-Hisp White")

#Outcome ~ Predictor + Covariates
model2_cost <- svyglm(Cost ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())
model2_struc <- svyglm(Structural ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())
model2_min <- svyglm(Minimization ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())
model2_lowp <- svyglm(Low_perc_eff ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())
model2_stigma <- svyglm(Stigma ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())

wb <- createWorkbook()

model_list <- list(
  Cost         = model2_cost,
  Structural   = model2_struc,
  Minimization = model2_min,
  Low_Perc_Eff = model2_lowp,
  Stigma       = model2_stigma
)

for (name in names(model_list)) {
  sheet_data <- tidy(model_list[[name]],
                     exponentiate = TRUE, conf.int = TRUE) %>%
    select(term, estimate, conf.low, conf.high, p.value) %>%
    rename(
      OR = estimate,
      `95% CI Lower` = conf.low,
      `95% CI Upper` = conf.high,
      `P-value` = p.value
    )
  
  addWorksheet(wb, name)
  writeData(wb, sheet = name, sheet_data)
}

saveWorkbook(wb, "model2_output2.xlsx", overwrite = TRUE)
```

```{r}
#Outcome ~ Predictor + Interaction
model3_cost <- svyglm(Cost ~ Distress * Race, design= NSDUH_subset, family=quasibinomial())
model3_struc <- svyglm(Structural ~ Distress * Race, design= NSDUH_subset, family=quasibinomial())
model3_min <- svyglm(Minimization ~ Distress * Race, design= NSDUH_subset, family=quasibinomial())
model3_lowp <- svyglm(Low_perc_eff ~ Distress * Race, design= NSDUH_subset, family=quasibinomial())
model3_stigma <- svyglm(Stigma ~ Distress * Race, design= NSDUH_subset, family=quasibinomial())

wb <- createWorkbook()

model_list <- list(
  Cost         = model3_cost,
  Structural   = model3_struc,
  Minimization = model3_min,
  Low_Perc_Eff = model3_lowp,
  Stigma       = model3_stigma
)

for (name in names(model_list)) {
  sheet_data <- tidy(model_list[[name]],
                     exponentiate = TRUE, conf.int = TRUE) %>%
    select(term, estimate, conf.low, conf.high, p.value) %>%
    rename(
      OR = estimate,
      `95% CI Lower` = conf.low,
      `95% CI Upper` = conf.high,
      `P-value` = p.value
    )
  
  addWorksheet(wb, name)
  writeData(wb, sheet = name, sheet_data)
}

saveWorkbook(wb, "model3_output.xlsx", overwrite = TRUE)
```

```{r}
#Outcome ~ Predictor + Covariates + Interaction
model4_cost <- svyglm(Cost ~ Distress * Race + Age + Sex + Marital_status + Education + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())
model4_struc <- svyglm(Structural ~ Distress * Race + Age + Sex + Marital_status + Education + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())
model4_min <- svyglm(Minimization ~ Distress * Race + Age + Sex + Marital_status + Education + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())
model4_lowp <- svyglm(Low_perc_eff ~ Distress * Race + Age + Sex + Marital_status + Education + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())
model4_stigma <- svyglm(Stigma ~ Distress * Race + Age + Sex + Marital_status + Education + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=quasibinomial())

wb <- createWorkbook()

model_list <- list(
  Cost         = model4_cost,
  Structural   = model4_struc,
  Minimization = model4_min,
  Low_Perc_Eff = model4_lowp,
  Stigma       = model4_stigma
)

for (name in names(model_list)) {
  sheet_data <- tidy(model_list[[name]],
                     exponentiate = TRUE, conf.int = TRUE) %>%
    select(term, estimate, conf.low, conf.high, p.value) %>%
    rename(
      OR = estimate,
      `95% CI Lower` = conf.low,
      `95% CI Upper` = conf.high,
      `P-value` = p.value
    )
  
  addWorksheet(wb, name)
  writeData(wb, sheet = name, sheet_data)
}

saveWorkbook(wb, "model4_output.xlsx", overwrite = TRUE)
```

# ROC curve comparability to gsem
```{r}
pdf("roc_plot4.pdf", width = 10, height = 8)

library(pROC)

logit_pred_cost4   <- predict(model4_cost, type = "response")
logit_pred_struct4 <- predict(model4_struc, type = "response")
logit_pred_min4    <- predict(model4_min, type = "response")
logit_pred_lowp4   <- predict(model4_lowp, type = "response")
logit_pred_stigma4 <- predict(model4_stigma, type = "response")


roc_cost4   <- roc(NSDUH_subset$variables$Cost, logit_pred_cost4, levels = c("0", "1"), direction = "<")
roc_struct4 <- roc(NSDUH_subset$variables$Structural, logit_pred_struct4, levels = c("0", "1"), direction = "<")
roc_min4    <- roc(NSDUH_subset$variables$Minimization, logit_pred_min4, levels = c("0", "1"), direction = "<")
roc_lowp4   <- roc(NSDUH_subset$variables$Low_perc_eff, logit_pred_lowp4, levels = c("0", "1"), direction = "<")
roc_stigma4 <- roc(NSDUH_subset$variables$Stigma, logit_pred_stigma4, levels = c("0", "1"), direction = "<")


auc_cost4   <- round(auc(roc_cost4), 3)
auc_struct4 <- round(auc(roc_struct4), 3)
auc_min4    <- round(auc(roc_min4), 3)
auc_lowp4   <- round(auc(roc_lowp4), 3)
auc_stigma4 <- round(auc(roc_stigma4), 3)

plot(roc_cost4, col = "red", lwd = 2, main = "ROC Curves for Logit Model4")
plot(roc_struct4, col = "blue", add = TRUE, lwd = 2)
plot(roc_min4, col = "green", add = TRUE, lwd = 2)
plot(roc_lowp4, col = "orange", add = TRUE, lwd = 2)
plot(roc_stigma4, col = "purple", add = TRUE, lwd = 2)

legend("bottomright", 
       legend = c(paste("Cost (AUC =", auc_cost4, ")"),
                  paste("Structural (AUC =", auc_struct4, ")"),
                  paste("Minimization (AUC =", auc_min4, ")"),
                  paste("Low Perc Eff (AUC =", auc_lowp4, ")"),
                  paste("Stigma (AUC =", auc_stigma4, ")")),
       col = c("red", "blue", "green", "orange", "purple"), 
       lwd = 2,
       bty="n", 
       inset=0)

roc_plot4 <- recordPlot()
```
```{r}
pdf("roc_plot2.pdf", width = 10, height = 8)

library(pROC)

logit_pred_cost2   <- predict(model2_cost, type = "response")
logit_pred_struct2 <- predict(model2_struc, type = "response")
logit_pred_min2    <- predict(model2_min, type = "response")
logit_pred_lowp2   <- predict(model2_lowp, type = "response")
logit_pred_stigma2 <- predict(model2_stigma, type = "response")

roc_cost2   <- roc(NSDUH_subset$variables$Cost, logit_pred_cost2, levels = c("0", "1"), direction = "<")
roc_struct2 <- roc(NSDUH_subset$variables$Structural, logit_pred_struct2, levels = c("0", "1"), direction = "<")
roc_min2    <- roc(NSDUH_subset$variables$Minimization, logit_pred_min2, levels = c("0", "1"), direction = "<")
roc_lowp2   <- roc(NSDUH_subset$variables$Low_perc_eff, logit_pred_lowp2, levels = c("0", "1"), direction = "<")
roc_stigma2 <- roc(NSDUH_subset$variables$Stigma, logit_pred_stigma2, levels = c("0", "1"), direction = "<")

auc_cost2   <- round(auc(roc_cost2), 3)
auc_struct2 <- round(auc(roc_struct2), 3)
auc_min2    <- round(auc(roc_min2), 3)
auc_lowp2   <- round(auc(roc_lowp2), 3)
auc_stigma2 <- round(auc(roc_stigma2), 3)

plot(roc_cost2, col = "red", lwd = 2, main = "ROC Curves for Logit Model2")
plot(roc_struct2, col = "blue", add = TRUE, lwd = 2)
plot(roc_min2, col = "green", add = TRUE, lwd = 2)
plot(roc_lowp2, col = "orange", add = TRUE, lwd = 2)
plot(roc_stigma2, col = "purple", add = TRUE, lwd = 2)

legend("bottomright", 
       legend = c(paste("Cost (AUC =", auc_cost2, ")"),
                  paste("Structural (AUC =", auc_struct2, ")"),
                  paste("Minimization (AUC =", auc_min2, ")"),
                  paste("Low Perc Eff (AUC =", auc_lowp2, ")"),
                  paste("Stigma (AUC =", auc_stigma2, ")")),
       col = c("red", "blue", "green", "orange", "purple"), 
       lwd = 2,
       bty="n",
       inset=0)

roc_plot2 <- recordPlot()
```

```{r}
pdf("roc_plot1.pdf", width = 10, height = 8)
library(pROC)

logit_pred_cost1   <- predict(model1_cost, type = "response")
logit_pred_struct1 <- predict(model1_struc, type = "response")
logit_pred_min1    <- predict(model1_min, type = "response")
logit_pred_lowp1   <- predict(model1_lowp, type = "response")
logit_pred_stigma1 <- predict(model1_stigma, type = "response")

roc_cost1   <- roc(NSDUH_subset$variables$Cost, logit_pred_cost1, levels = c("0", "1"), direction = "<")
roc_struct1 <- roc(NSDUH_subset$variables$Structural, logit_pred_struct1, levels = c("0", "1"), direction = "<")
roc_min1    <- roc(NSDUH_subset$variables$Minimization, logit_pred_min1, levels = c("0", "1"), direction = "<")
roc_lowp1   <- roc(NSDUH_subset$variables$Low_perc_eff, logit_pred_lowp1, levels = c("0", "1"), direction = "<")
roc_stigma1 <- roc(NSDUH_subset$variables$Stigma, logit_pred_stigma1, levels = c("0", "1"), direction = "<")

auc_cost1   <- round(auc(roc_cost1), 3)
auc_struct1 <- round(auc(roc_struct1), 3)
auc_min1    <- round(auc(roc_min1), 3)
auc_lowp1   <- round(auc(roc_lowp1), 3)
auc_stigma1 <- round(auc(roc_stigma1), 3)

plot(roc_cost1, col = "red", lwd = 2, main = "ROC Curves for Logit Model1")
plot(roc_struct1, col = "blue", add = TRUE, lwd = 2)
plot(roc_min1, col = "green", add = TRUE, lwd = 2)
plot(roc_lowp1, col = "orange", add = TRUE, lwd = 2)
plot(roc_stigma1, col = "purple", add = TRUE, lwd = 2)

legend("bottomright", 
       legend = c(paste("Cost (AUC =", auc_cost1, ")"),
                  paste("Structural (AUC =", auc_struct1, ")"),
                  paste("Minimization (AUC =", auc_min1, ")"),
                  paste("Low Perc Eff (AUC =", auc_lowp1, ")"),
                  paste("Stigma (AUC =", auc_stigma1, ")")),
       col = c("red", "blue", "green", "orange", "purple"), 
       lwd = 2,
       bty= "n",
       inset=0)

roc_plot1 <- recordPlot()
```

```{r}
pdf("roc_plot3.pdf", width = 10, height = 8)

library(pROC)

logit_pred_cost3   <- predict(model3_cost, type = "response")
logit_pred_struct3 <- predict(model3_struc, type = "response")
logit_pred_min3    <- predict(model3_min, type = "response")
logit_pred_lowp3   <- predict(model3_lowp, type = "response")
logit_pred_stigma3 <- predict(model3_stigma, type = "response")

roc_cost3   <- roc(NSDUH_subset$variables$Cost, logit_pred_cost3, levels = c("0", "1"), direction = "<")
roc_struct3 <- roc(NSDUH_subset$variables$Structural, logit_pred_struct3, levels = c("0", "1"), direction = "<")
roc_min3    <- roc(NSDUH_subset$variables$Minimization, logit_pred_min3, levels = c("0", "1"), direction = "<")
roc_lowp3   <- roc(NSDUH_subset$variables$Low_perc_eff, logit_pred_lowp3, levels = c("0", "1"), direction = "<")
roc_stigma3 <- roc(NSDUH_subset$variables$Stigma, logit_pred_stigma3, levels = c("0", "1"), direction = "<")

auc_cost3   <- round(auc(roc_cost3), 3)
auc_struct3 <- round(auc(roc_struct3), 3)
auc_min3    <- round(auc(roc_min3), 3)
auc_lowp3   <- round(auc(roc_lowp3), 3)
auc_stigma3 <- round(auc(roc_stigma3), 3)

plot(roc_cost3, col = "red", lwd = 2, main = "ROC Curves for Logit Model3")
plot(roc_struct3, col = "blue", add = TRUE, lwd = 2)
plot(roc_min3, col = "green", add = TRUE, lwd = 2)
plot(roc_lowp3, col = "orange", add = TRUE, lwd = 2)
plot(roc_stigma3, col = "purple", add = TRUE, lwd = 2)

legend("bottomright", 
       legend = c(paste("Cost (AUC =", auc_cost3, ")"),
                  paste("Structural (AUC =", auc_struct3, ")"),
                  paste("Minimization (AUC =", auc_min3, ")"),
                  paste("Low Perc Eff (AUC =", auc_lowp3, ")"),
                  paste("Stigma (AUC =", auc_stigma3, ")")),
       col = c("red", "blue", "green", "orange", "purple"), 
       lwd = 2,
       bty = "n",      
       inset = 0)

roc_plot3 <- recordPlot()
```

# AUC Plot
```{r}
pdf("roc_faceted_plot.pdf", width = 10, height = 8)

par(mfrow = c(2, 2),           
    oma = c(0, 0, 3, 0))       

replayPlot(roc_plot1)

replayPlot(roc_plot2)

replayPlot(roc_plot3)

replayPlot(roc_plot4)


dev.off()
```



```{r}
library(ggalt)  # if you want dumbbell style

auc_wide <- auc %>%
  pivot_wider(names_from = Link, values_from = AUC)

auc_wide <- auc_wide %>%
  mutate(overlap = Logit == Probit)

p <- ggplot(auc_wide, aes(x = Logit, xend = Probit, y = Barrier)) +
  geom_dumbbell(
    color = "gray80",
    size = 3,
    colour_x = "blue",
    colour_xend = "red"
  ) +
  # Add a subtle mark for overlapping AUCs
  geom_point(data = filter(auc_wide, overlap),
             aes(x = Logit, y = Barrier),
             color = "purple", size = 3, shape = 4) +  # X shape
  facet_wrap(~ Model) + 
  scale_color_manual(
    name = "Link Function",
    values = c("Logit" = "blue", "Probit" = "red")
  ) +
  labs(
    title = "Logit vs. Probit AUCs by Barrier and Model",
    x = "AUC",
    y = "Barrier",
    caption = "Purple '×' indicates Logit = Probit; Logit= Blue; Probit= Red"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")

ggsave("AUC_dumbbell_plot.png", plot = p, width = 10, height = 6, dpi = 300)

auc_logit <- auc %>% filter(Link == "Logit") %>% select(!label)
```


```{r}
auc_logit <- auc %>% filter(Link == "Logit") %>% select(!label)

plot <- ggplot(auc_logit, aes(x = AUC, y = Barrier)) + geom_point(color = "red", size = 3) + facet_wrap(~Model) + theme_bw()

ggsave("AUC_plot.png", plot = plot, width = 10, height = 6, dpi = 300)
```

# Stratified Analysis Repeat
```{r}
race <- c("Non-Hisp White", "Non-Hisp Black/Afr Am", "Non-Hisp Asian", "Hispanic", "Non-Hisp Multi Race")
barriers <- c("Cost", "Structural", "Minimization", "Low_perc_eff", "Stigma")

for (race in race) {
  cat("\nRace:", race, "\n")
  subset_design <- subset(NSDUH_subset, Race == race)
  print(table(subset_design$variables$Distress))
}
```


```{r}
models_by_race <- list()

#Cost 
for (this_race in race) {  
  
  subset_design <- subset(NSDUH_subset, Race == this_race)
  
  if (length(unique(subset_design$variables$Cost)) > 1) {
    
    model <- svyglm(
      Cost ~ Distress,
      design = subset_design,
      family = quasibinomial()
    )
    
    models_by_race[[this_race]] <- model
    
    cat("\n=== Model for", this_race, "===\n")
    print(summary(model))
    
  } else {
    
    cat("\n=== Skipped", this_race, "- Cost has only one level ===\n")
    models_by_race[[this_race]] <- NA
  }
}

#Structural
models_by_race <- list()

for (this_race in race) {  # <-- use 'this_race' for the current race
  
  subset_design <- subset(NSDUH_subset, Race == this_race)
  
  if (length(unique(subset_design$variables$Cost)) > 1) {
    
    model <- svyglm(
      Structural ~ Distress,
      design = subset_design,
      family = quasibinomial()
    )
    
    models_by_race[[this_race]] <- model
    
    cat("\n=== Model for", this_race, "===\n")
    print(summary(model))
    
  } else {
    
    cat("\n=== Skipped", this_race, "- Structural has only one level ===\n")
    models_by_race[[this_race]] <- NA
  }
}

# Minimization
models_by_race <- list()

for (this_race in race) {  # <-- use 'this_race' for the current race
  
  subset_design <- subset(NSDUH_subset, Race == this_race)
  
  if (length(unique(subset_design$variables$Cost)) > 1) {
    
    model <- svyglm(
      Minimization ~ Distress,
      design = subset_design,
      family = quasibinomial()
    )
    
    models_by_race[[this_race]] <- model
    
    cat("\n=== Model for", this_race, "===\n")
    print(summary(model))
    
  } else {
    
    cat("\n=== Skipped", this_race, "- Minimization has only one level ===\n")
    models_by_race[[this_race]] <- NA
  }
}

#Low perc 
models_by_race <- list()

for (this_race in race) {  # <-- use 'this_race' for the current race
  
  subset_design <- subset(NSDUH_subset, Race == this_race)
  
  if (length(unique(subset_design$variables$Cost)) > 1) {
    
    model <- svyglm(
      Low_perc_eff ~ Distress,
      design = subset_design,
      family = quasibinomial()
    )
    
    models_by_race[[this_race]] <- model
    
    cat("\n=== Model for", this_race, "===\n")
    print(summary(model))
    
  } else {
    
    cat("\n=== Skipped", this_race, "- Low_perc_eff has only one level ===\n")
    models_by_race[[this_race]] <- NA
  }
}

#Stigma 
models_by_race <- list()

for (this_race in race) {  # <-- use 'this_race' for the current race
  
  subset_design <- subset(NSDUH_subset, Race == this_race)
  
  if (length(unique(subset_design$variables$Cost)) > 1) {
    
    model <- svyglm(
      Stigma ~ Distress,
      design = subset_design,
      family = quasibinomial()
    )
    
    models_by_race[[this_race]] <- model
    
    cat("\n=== Model for", this_race, "===\n")
    print(summary(model))
    
  } else {
    
    cat("\n=== Skipped", this_race, "- Stigma has only one level ===\n")
    models_by_race[[this_race]] <- NA
  }
}
```

```{r}
library(broom)

barriers <- c("Cost", "Structural", "Minimization", "Low_perc_eff", "Stigma")

races <- c("Non-Hisp White", "Non-Hisp Black/Afr Am", "Non-Hisp Asian", "Hispanic", "Non-Hisp Multi Race")


all_results <- data.frame()


for (barrier in barriers) {
  
  cat("\n=== Running models for barrier:", barrier, "===\n")
  
  for (this_race in races) {
    
    subset_design <- subset(NSDUH_subset, Race == this_race)
    
    if (length(unique(subset_design$variables[[barrier]])) > 1) {
      
      model <- svyglm(
        as.formula(paste0(barrier, " ~ Distress")),
        design = subset_design,
        family = quasibinomial()
      )
      
      tidy_model <- tidy(model, conf.int = TRUE, exponentiate = TRUE)
      
      tidy_model <- tidy_model[grepl("Distress", tidy_model$term), ]
      
      tidy_model$Race <- this_race
      tidy_model$Barrier <- barrier
      
      all_results <- rbind(all_results, tidy_model)
      
    } else {
      
      cat("=== Skipped", this_race, "-", barrier, "- no variability ===\n")
      
    }
  }
}

```

# Export 
```{r}
all_results$estimate <- round(all_results$estimate, 3)
all_results$conf.low <- round(all_results$conf.low, 3)
all_results$conf.high <- round(all_results$conf.high, 3)
all_results$p.value <- round(all_results$p.value, 3)

library(writexl)

write_xlsx(all_results, "all_results.xlsx")

```

```{r}

#Outcome ~ Predictor + Covariates
model2_cost <- svyglm(Cost ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=binomial())
model2_struc <- svyglm(Structural ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=binomial())
model2_min <- svyglm(Minimization ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=binomial())
model2_lowp <- svyglm(Low_perc_eff ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=binomial())
model2_stigma <- svyglm(Stigma ~ Distress + Age + Sex + Marital_status + Education + Race + Urbanicity + Poverty + Health_status + Employment_status + SUD + Insurance + Disability + Income + Drug_Use + Alcohol_Use, design= NSDUH_subset, family=binomial())

wb <- createWorkbook()

model_list <- list(
  Cost         = model2_cost,
  Structural   = model2_struc,
  Minimization = model2_min,
  Low_Perc_Eff = model2_lowp,
  Stigma       = model2_stigma
)

for (name in names(model_list)) {
  sheet_data <- tidy(model_list[[name]],
                     exponentiate = TRUE, conf.int = TRUE) %>%
    select(term, estimate, conf.low, conf.high, p.value) %>%
    rename(
      OR = estimate,
      `95% CI Lower` = conf.low,
      `95% CI Upper` = conf.high,
      `P-value` = p.value
    )
  
  addWorksheet(wb, name)
  writeData(wb, sheet = name, sheet_data)
}

saveWorkbook(wb, "model2_binomial.xlsx", overwrite = TRUE)

library(pROC)

logit_pred_cost   <- predict(model2_cost, type = "response")
logit_pred_struct <- predict(model2_struc, type = "response")
logit_pred_min    <- predict(model2_min, type = "response")
logit_pred_lowp   <- predict(model2_lowp, type = "response")
logit_pred_stigma <- predict(model2_stigma, type = "response")

roc_cost   <- roc(model2_cost$y, logit_pred_cost)
roc_struct <- roc(model2_struc$y, logit_pred_struct)
roc_min    <- roc(model2_min$y, logit_pred_min)
roc_lowp   <- roc(model2_lowp$y, logit_pred_lowp)
roc_stigma <- roc(model2_stigma$y, logit_pred_stigma)

auc_cost   <- round(auc(roc_cost), 3)
auc_struct <- round(auc(roc_struct), 3)
auc_min    <- round(auc(roc_min), 3)
auc_lowp   <- round(auc(roc_lowp), 3)
auc_stigma <- round(auc(roc_stigma), 3)

plot(roc_cost, col = "red", lwd = 2, main = "ROC Curves for Logit Model2 Binomial")

plot(roc_struct, col = "blue", add = TRUE, lwd = 2)
plot(roc_min, col = "green", add = TRUE, lwd = 2)
plot(roc_lowp, col = "orange", add = TRUE, lwd = 2)
plot(roc_stigma, col = "purple", add = TRUE, lwd = 2)

legend("bottomright", 
       legend = c(paste("Cost (AUC =", auc_cost, ")"),
                  paste("Structural (AUC =", auc_struct, ")"),
                  paste("Minimization (AUC =", auc_min, ")"),
                  paste("Low Perc Eff (AUC =", auc_lowp, ")"),
                  paste("Stigma (AUC =", auc_stigma, ")")),
       col = c("red", "blue", "green", "orange", "purple"), 
       lwd = 2)
```



